---
title: "Problem Set 7"
author: "Westley Cook"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# standard first load

library(tidyverse)

# for use of skim()

library(skimr)

# for using tidy()

library(broom)

# for making tables

library(gt)

# loading and naming congress data

nominate <- read_csv("raw-data/HSall_members.csv", col_types = cols())

```

## Question 1: Exploratory Data Analysis

### 1A) Summary Statistics

```{r question_1a, echo=FALSE}

# This r chunk wrangles the data as requested in the assignment prompt, giving a
# new name to the object for easy use in parts b) and c), then prints the
# results of the skim() function to show some summary stats

# I filter first for just the senate and house, using the vertical bar for "or",
# and additionally filter for just the 116th congress. I use case_when to assign
# names to the variable party, and then filter again to drop the resulting NA
# values (for those neither democrat nor republican, the independents). I then
# use percent_rank() to add percentile values based on nominate_dim1,
# multiplying by 100 to comply with the assignment prompt formatting statement
# (that the median value should be 50)

nominate_116 <- nominate %>% 
  filter(chamber == "Senate" | chamber == "House",
         congress == 116) %>% 
  mutate(party = case_when(party_code == 100 ~ "Democrat",
                           party_code == 200 ~ "Republican"),
         age = 2019 - born) %>% 
  filter(!is.na(party)) %>% 
  mutate(nominate_percentile = percent_rank(nominate_dim1) * 100)

# Having wrangled the data, I select the two variables of interest, pipe it to
# the skim() function, and print the result

nominate_116 %>% 
  select(nominate_percentile, age) %>% 
  skim()

```

### 1B) Visualizations

```{r question_1b, echo=FALSE}

# This r chunk creates two plots

# The first plot shows the distribution of age in congress by chamber. I found
# this data to be interesting in that there are more young representatives than
# young senators. I think this plot does a nice job preserving a clean look at
# the overall distribution of ages while adding the insight that there's more
# (relative) youth in the House. I chose brown and black for the colors because
# I wanted something distinctly nonpartisan; I didn't want there to by any
# confusion about party affiliations by choosing something like, say, red and
# blue to show the different chambers of congress

nominate_116 %>% 
  ggplot(aes(age, fill = chamber)) +
  geom_bar() +
  theme_classic() +
  labs(title = "Distribution of Congressional Ages by Chamber",
       subtitle = "More young Representatives than young Senators",
       x = "Age",
       y = "Count") +
  scale_fill_manual(name = "Chamber",
                    values = c("House" = "brown",
                               "Senate" = "black"))

# The second plot shows the relationship between age and party. I wasn't sure if
# I should make this plot, given that the prompt asked us to make one for
# nominate_percentile, but a histogram of percentile rank is essentially flat
# (given the nature of percentiles; they're spread out evenly, as seen in the
# skim() results) so that seemed really boring. So I thought a different look at
# ages might be warranted. Using a combination of fill = party and facet_wrap(~
# party) and putting the plots above/below one another seemed to give the
# cleanest look at the differences in distribution. A single bar plot with
# position = "stack" was less clear, and position = "dodge" generated really
# skinny bars that also made it less clear. Used legend.position = "none" to
# take away the legend, which facet_wrap() renders unnecessary

nominate_116 %>% 
  ggplot(aes(age, fill = party)) +
  geom_bar() +
  facet_wrap(~ party, ncol = 1) +
  theme_classic() +
  labs(title = "Distribution of Congressional Ages by Party",
       subtitle = "Slightly wider age range for Democrats than for Republicans",
       x = "Age",
       y = "Count") +
  scale_fill_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  theme(legend.position = "none")

```

### 1C) Bivariate Correlations

```{r question 1c, echo=FALSE}

# This r chunk finds the value of the correlation coefficient for age and
# nominate_percentile and assigns that value to an object to insert below. I
# round to 3 digits arbitrarily because 2 seemed too few and 3 seemed too many

cor_coeff <- nominate_116 %>% 
  summarize(cor = cor(age, nominate_percentile)) %>% 
  pull(cor) %>% 
  round(digits = 3)

```

The correlation coefficient for age and nominate_percentile for the 116th Congress is **`r cor_coeff`**.

The negative coefficient indicates that the variables are moving in *inverse* directions; as age goes up, the percentile rank of ideology goes down. If the coefficient were positive, age and percentile rank would move in the same direction.

### 1D) Plotting Bivariate Relationships

```{r question_1d, echo=FALSE}

# This r chunk creates a bivariate plot of age and nominate_percentile

# Eyeballing the scatter plot before adding the regression lines, it didn't look
# like there was much of a relationship between age and percentile, but adding
# geom_smooth() revealed the negative association (which does, in fact, match
# the sign of the correlation coefficient). I got the aggregate line to show up
# on the plot by taking aes(color = party) out of the ggplot layer and adding a
# geom_smooth() layer; then I got the colored points and lines separated by
# party to show up by adding the color mapping to the geom_point layer and to
# the second geom_smooth layer. It was then pretty simple to add labels and
# manually change colors to produce the final result

# Note: the order in my legend is flipped from the plot we're replicating
# (Republicans are first here but not there) because Republican points are on
# top in the plot itself and I thought keeping that consistent in the legend
# made for a nicer-looking plot

nominate_116 %>% 
  ggplot(aes(age, nominate_percentile)) +
  geom_smooth(method = "glm", se = FALSE, color = "black") +
  geom_point(aes(color = party)) +
  geom_smooth(aes(color = party), method = "glm", se = FALSE) +
  theme_classic() +
  labs(title = "Relationship Between Age and DW Nominate Percentile",
       subtitle = "Higher scores represent more conservative positions",
       x = "Age",
       y = "Percentile Rank (DW Nominate)") +
  scale_color_manual(name = "Party",
                     breaks = c("Republican", "Democrat"),
                     values = c("Democrat" = "blue", "Republican" = "red"))

```

## Question 2: Single Regression

### 2A) Using lm()

```{r question_2a, echo=FALSE}

# This r chunk runs a single-variable linear regression using data for Democrats
# in the 116th Congress

# Using the modified data set from question 1, I first filter for just the
# Democrats. Then I use lm() to model the effect of age on nominate_percentile
# and tidy() with conf.int = TRUE to produce a clean tibble with the upper and
# lower bounds. I then select the variables of interest and pipe them to a gt()
# table. Finally, I add a title, subtitle, and column labels

nominate_116 %>% 
  filter(party == "Democrat") %>% 
  lm(nominate_percentile ~ age, data = .) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, conf.low, conf.high) %>% 
  gt() %>% 
  tab_header(title = "Effect of Age on DW Nominate Percentile",
             subtitle = "Higher percentage suggests more conservative") %>% 
  cols_label(term = "Variable",
             estimate = "Estimate",
             conf.low = "Lower bound",
             conf.high = "Upper bound")

```

### 2B) Interpreting results

The estimated average treatment effect is that being one year older is associated, on average, with a decrease of about one-third of a percentile (-0.33 percentile units) for a Democratic member of Congress. We should not necessarily interpret this correlation as being causal unless we’re certain that no other variables could be confounding the result (an assumption which seems unlikely). For example, perhaps it’s not age itself that causes Congress members to become more liberal, but rather it’s that older members of Congress aren’t as frequently trying to be re-elected and thus don’t need to pander to the median voter by feigning conservatism.

We calculate a confidence interval around our average treatment effect because we don’t think that our sample of Democratic Congress members in the 116th Congress is necessarily representative of all Democratic Congress members in other sessions; calculating a confidence interval allows us to have a plausible range of values for the treatment effect (as low as -0.469 or as high as -0.196) if we want to extrapolate the correlation we found in this particular data to past or future sessions of Congress.

### 2C) Regression and the Rubin Causal Model

Because treatment was randomly assigned, this analysis could be considered causal. The slope coefficient X on military would then mean that the *average treatment effect* on nominate_percentile of having a Congress member serve in the military would be X units. The average treatment effect would represent the average difference in potential outcomes (comparing potential percentile rank *with* military service and potential percentile rank *without* military service) for Democratic members of Congress.

## Question 3: Many Regressions

```{r question_3, echo=FALSE}

# This r chunk replicates the assignment's plot of average treatment effect of
# age on conservatism among democrats over time

dems_nominate <- nominate %>% 
  filter(chamber == "Senate" | chamber == "House",
         party_code == 100) %>% 
  mutate(nominate_percentile = percent_rank(nominate_dim1) * 100,
         start_date = congress + 1812 + (congress - 25),
         age = start_date - born) %>% 
  group_by(congress) %>% 
  nest() %>% 
  mutate(mod = map(data, ~ lm(nominate_percentile ~ age, data = .)),
         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
         coef = map_dbl(reg_results, ~ filter(., term == "age") %>% 
                          pull(estimate)),
         upper = map_dbl(reg_results, ~ filter(., term == "age") %>% 
                           pull(conf.high)),
         lower = map_dbl(reg_results, ~ filter(., term == "age") %>% 
                           pull(conf.low)))

dems_nominate %>% 
  ggplot(aes(congress, coef)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(x = congress, ymin = lower, ymax = upper), color = "blue") +
  theme_classic() +
  labs(title = "Average Treatment Effect of Age on Conservatism 
 Among Democrats Over Time",
       subtitle = "Scores use DW Nominate Dimension 1
 95% Confidence Interval",
       x = "Congress",
       y = "Estimate")

```

##### Worked with:
None







